---
title: "Carvalhoetal_Bdmodel_statisticalanalyses"
author: "Tamilie Carvalho"
date: "2023-03-11"
output: html_document
---

Loading required packages
```{r}
library(survival)
library(survminer)
library(ggplot2)
```


**Bd dosage response experiment**

To import the data file
```{r}
d<-data.frame(read.csv("Bd_dosage_response.csv"))
table(d$Treatment)
```

To compare survival curves across treatments
```{r}
# To create a survival object.
surv1<-Surv(time = d$Mortality_day, event = d$Censor)

# To fit a survival curves.
surv2<-survfit(surv1 ~ Treatment, data = d)

summary(surv2)$table
```

To compute a log-rank test.
```{r}
surv3<-survdiff(Surv(Mortality_day, Censor) ~ Treatment, data = d)
surv3

## To conduct pairwise comparisons.
surv4<-pairwise_survdiff(Surv(Mortality_day, Censor)~Treatment, data = d)
surv4
```


To plot survival curves.
```{r}
p <- ggsurvplot(surv2, data = d, 
                 pval = TRUE, pval.size= 3.5, pval.coord = c(0.01, 0.01), 
                 palette = c("#2CD9F0","#2CD9F0", "#756E6D","#D200E6","#D200E6"),
                 pval.method = FALSE, surv.median.line = NULL, conf.int = FALSE, 
                 legend = c(0.09, 0.26), font.legend = c(8),
                 legend.labs=c("CLFT073_HD","CLFT073_LD","Control","JEL423_HD","JEL423_LD"),
                 legend.title= "Treatment", font.x=c(12), font.y=c(12),
                 font.tickslab=10, 
                 linetype = c("strata"))+
                 xlab("Days post-infection")
         

p<-p$plot + theme(plot.title = element_text(hjust = 0.5, size = 10))+
            scale_linetype_manual(values = c(1,2,1,1,2))+
            theme(axis.ticks =    element_line(size = 0.10), 
            axis.ticks.length = unit(0.10, "cm"))
p

ggsave("Bddosage.png", p, height = 5, width = 7.5, dpi = 600)
```


To clean R
```{r}
a<-objects()
rm(list = a)
```

**Model generalizability experiment**

To import the data file.
```{r}
d<-data.frame(read.csv("Modelgeneralizability.csv"))

table(d$Genotype)
```


To remove control group.
```{r}
d1<-data.frame(subset(d, Genotype=="CLFT137" | Genotype== "CLFT150" | Genotype=="CLFT142" | Genotype=="CLFT024-02" | Genotype=="CLFT156" | Genotype=="CLFT039")) 

d1<-droplevels(d1) 

table(d1$Genotype)
```

To create a survival object.
```{r}
surv1<-Surv(time = d1$Mortality_day, event = d1$Censor)

# To fit a survival curves for genotypes.
surv2g<-survfit(surv1 ~ Genotype, data = d1)
summary(surv2g)$table

surv3g<-survdiff(Surv(Mortality_day, Censor) ~ Genotype, data = d1)
surv3g
```


To conduct pairwise comparisons across genotypes.
```{r}
surv4g<-pairwise_survdiff(Surv(Mortality_day, Censor)~Genotype, data = d1)
surv4g
```


To plot survival curves for genotypes.
```{r}
p <- ggsurvplot(surv2g, data = d1, 
                 pval = TRUE, pval.size= 3.5, pval.coord = c(0.03, 0.01), 
                 palette = c("#DECD18","#DECD18","#1819E0","#019302","#019302","#1819E0"),
                 pval.method = FALSE, surv.median.line = NULL, conf.int = FALSE, 
                 legend = c(0.12, 0.34), font.legend = c(8),
                 legend.labs=c("CLFT039","CLFT024-02","CLFT137","CLFT142","CLFT150", "CLFT156"),
                 legend.title= "Genotype", font.x=c(12), font.y=c(12),
                 font.tickslab=10, 
                 linetype = c("strata"))+
                 xlab("Days post-infection")
         

p<-p$plot + theme(plot.title = element_text(hjust = 0.5, size = 10))+
            scale_linetype_manual(values = c(1,2,1,1,2,2))+
            theme(axis.ticks =    element_line(size = 0.10), 
            axis.ticks.length = unit(0.10, "cm"))
p

ggsave("Genotypes2.png", p, height = 4, width = 7.5, dpi = 600)

```



To fit a survival curves for lineages
```{r}
surv2l<-survfit(surv1 ~ Lineage, data = d1)
summary(surv2l)$table

surv3l<-survdiff(Surv(Mortality_day, Censor) ~ Lineage, data = d1)
surv3l
```


To conduct pairwise comparisons across lineages.
```{r}
surv4l<-pairwise_survdiff(Surv(Mortality_day, Censor)~Lineage, data = d1)
surv4l
```

To plot survival curves for lineages.
```{r}

p <- ggsurvplot(surv2l, data = d1,
            pval = TRUE, pval.size= 3.5, pval.coord = c(0.01, 0.01),
            palette = c("#019302","#1819E0", "#DECD18"),
            pval.method = FALSE, surv.median.line = NULL, 
            legend = c(0.12, 0.34),font.legend = c(8), 
            legend.title= "Lineage",legend.labs=c("Bd-Asia-2/Brazil","Bd-GPL","Hybrid"), 
            font.tickslab=10,
            conf.int = FALSE, font.x=c(12), font.y=c(12))+
            xlab("Days post-infection")


p<-p$plot + theme(plot.title = element_text(hjust = 0.5, size = 10))+
            theme(axis.ticks =    element_line(size = 0.10), 
            axis.ticks.length = unit(0.10, "cm"))
p

ggsave("Lineage.png", p, height = 4, width = 7.5, dpi = 600)
```



To clean R
```{r}
a<-objects()
rm(list = a)
```


**validation experiment**

Part 1 - Survival curves - To compare survival curves  results between model generalizability and validation experiments

To import the data file.
```{r}
dve<-data.frame(read.csv("Validation.csv"))
table(dve$Genotype_experiment)
```


To remove control group.
```{r}
dve<-data.frame(subset(dve, Genotype_experiment=="CLFT150_MG" | Genotype_experiment=="CLFT039_MG"  | Genotype_experiment=="CLFT150_V"  | Genotype_experiment=="CLFT039_V")) 

dve<-droplevels(dve) 

table(dve$Genotype_experiment)
````


To compare survival curves across treatments (genotypes/exp)
```{r}
# To create a survival object.
surv1<-Surv(time = dve$Mortality_day, event = dve$Censor)

# To fit a survival curves.
surv2ve<-survfit(surv1 ~ Genotype_experiment, data = dve)

summary(surv2ve)$table
```

To compute a log-rank test.
```{r}
surv3<-survdiff(Surv(Mortality_day, Censor) ~ Genotype_experiment, data = dve)
surv3

## To conduct pairwise comparisons.
surv4<-pairwise_survdiff(Surv(Mortality_day, Censor)~Genotype_experiment, data = dve)
surv4
```


To plot survival curves.
```{r}
p <- ggsurvplot(surv2ve, data = dve, 
                 pval = TRUE, pval.size= 3.5, pval.coord = c(0.03, 0.01), 
                 palette = c("#DECD18","#DECD18","#019302","#019302"),
                 pval.method = FALSE, surv.median.line = NULL, conf.int = FALSE, 
                 legend = c(0.12, 0.34), font.legend = c(8),
                 legend.labs=c("CLFT039_MG","CLFT039_V","CLFT150_MG","CLFT150_V"),
                 legend.title= "Genotype", font.x=c(12), font.y=c(12),
                 font.tickslab=10, 
                 linetype = c("strata"))+
                 xlab("Days post-infection")
         

p<-p$plot + theme(plot.title = element_text(hjust = 0.5, size = 10))+
            scale_linetype_manual(values = c(1,2,1,2))+
            theme(axis.ticks =    element_line(size = 0.10), 
            axis.ticks.length = unit(0.10, "cm"))
p

ggsave("EXP1+EXP2.png", p, height = 4, width = 7.5, dpi = 600)
````


Part 2 - Bd Load - To compare Bd load results between model generalizability and validation experiments

```{r}
library(RVAideMemoire) 
library(ggplot2)
library(multcomp)
library(rstatix)
library(car)
```

Data normality verification  
```{r}
dve$Genotype_experiment<-as.factor(dve$Genotype_experiment)
dve$Lineage<-as.factor(dve$Lineage)
dve["LogBdLoad"]<-c(log10(dve$Bd_Load))
str(dve)
head(dve)

shapiro_test(dve$LogBdLoad)
byf.shapiro(LogBdLoad ~ Genotype_experiment, dve) #Shapiro per group
leveneTest(LogBdLoad ~ Genotype_experiment, dve, center=mean)
boxplot(LogBdLoad ~ Genotype_experiment, data = dve, ylab="Bd_load", xlab="Genotype")
```

To run anova
```{r}
anova_Bdload <- aov(LogBdLoad ~ Genotype_experiment, dve)
summary(anova_Bdload)
plot(anova_Bdload)

# To compute a posthoc test 
posthocBdload<-glht(anova_Bdload, linfct=mcp(Genotype_experiment="Tukey"))
summary(posthocBdload)
```

To plot Boxplot
```{r}
cols <- c( "#019302","#DECD18")

g<-ggplot(dve, aes(x = Genotype_experiment, y = Bd_Load, fill = Lineage)) + 
  stat_boxplot(geom = "errorbar",
               width = 0.25) + 
  geom_boxplot(alpha = 1,          # Fill transparency
               colour = "#000000",   # Border color
               outlier.colour = 1) + # Outlier color
  scale_fill_manual(values = cols) +
  theme(axis.ticks =  element_line(size = 0.10), 
      axis.ticks.length = unit(0.10, "cm"))

boxplot<-g + theme_classic()

boxplot

ggsave("boxplot.png", boxplot, height = 4, width = 7.5, dpi = 600)
```







